---
title: "Measuring sentiment in Tuskegee Airmen web profiles"
date: 2022-02-08
output: html_output
---

# TidyTuesday

For this week's Tidy Tuesday, I wanted to take advantage of two cool features of this dataset: The geographic location (home state of pilots) and the web profiles available for select pilots. Looking through some of the profiles, I noticed that the experiences of the airmen described in the profiles varied drastically, and it would be neat to capture that variation somehow.  As such, I decided to 1) scrape the text from each available profile, 2) calculate a sentiment score for each profile, and then 3) compare the average sentiment across states.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

## install pacman package if not already installed.
if(!require(pacman)){
  install.packages("pacman")
}

# devtools::install_github("vladmedenica/themedubois") ## install themedbuois package from github 

pacman::p_load(tidyverse,tidytuesdayR,tidytext,scales,datasets,themedubois)

```

# Loading data

```{r Load}

tt <- tt_load("2022-02-08")

data <- tt$airmen ##extracting data frame

head(data)

```

# Analyzing web profiles

First I'm going to see how to access correct information in web profiles. First need to limit data to observations that have web profiles. Then, I will test out the first observation to see how to correctly parse relevant text.

```{r}
library(rvest)

## first grabbing first website to test
site_test <-  data %>%
  filter(!is.na(web_profile)) %>%
    select(web_profile) %>%
    slice(1) %>%
    pull()

## after poking around on the site with SelectorGadget, looks like the relevant html nodes is .elementor-widget-theme-post-content p. It does grab some extra text outside of the bios, but not sure how to fix this issue, given the site structure.

site_test %>%
  read_html() %>%
  html_nodes(".elementor-widget-theme-post-content p") %>%
  html_text2() %>%
  str_c(collapse = "")
```

Nice! Assuming the structure of pages is consistent, I now know how to access the bios on each wage. However, there is some extra text tagged on the end that I don't want. I'm going to leave it in for now, as I don't think the words included in the extra text will affect sentiment too much.

Now I need to iterate the process for each available web profile.


```{r}

## creating subset of data to only include rows with web pages
with_page <- data %>%
  filter(!is.na(web_profile)) 

## first going to create helper function which I will use to iterate over each row in data
get_text <- function(x) {
  read_html(x) %>%
  html_nodes(".elementor-widget-theme-post-content p") %>%
  html_text2() %>%
  str_c(collapse = "")
}

## now using map() to apply scraping function to each site

bios <- with_page %>%
  mutate(bio = map(web_profile, ~ get_text(.x)))
  

bios %>%
  select(bio) %>%
  slice(5) %>%
  pull()
```

# Analyze text

First I will tokenize bios into a tidytext format, with one word per row. I will then perform some typical text preprocessing, then calculate a sentiment score for each profile.

```{r}

tokens <-
  bios %>%
  unnest_tokens("word","bio", remove)

sentiment_scores <- tokens %>%
  anti_join(stop_words) %>%
  inner_join(sentiments) %>%
  count(name,sentiment) %>%
    pivot_wider(names_from = "sentiment",
              values_from = n,
              values_fill = 0) %>%
  mutate(positive_index = round(positive/(positive + negative),2)) ## measuring sentiment as proportion of positive/positive+negative words


## now combining back with original data 

data_combined <- with_page %>%
  inner_join(sentiment_scores) 
  

```

# Finding patterns of sentiment in profiles

I could plot sentiment scores as either continuous (heatmap) or categorical. 
Going with categorical to better implmeent DuBois colors. 

```{r}

## first need to grab full names of states for creating map later
state_info <- tibble("state" = state.abb, "state_name" = state.name)


state_data <- data_combined %>%
    group_by(state) %>%
  summarize(
    positive = mean(positive_index) ##taking average sentiment by state
  ) %>%
  right_join(state_info) %>%
  select("state" = state_name, positive) %>%
  mutate(state = tolower(state),
         index = scales::rescale(positive), ## rescaling sentiment index for possible continuous outcome
         terciles = factor(ntile(index,3), ## grouping index by terciles for categorical version
                       levels = c(1, 2, 3),
                       labels = c("Negative", "Neutral", "Positive"))
         )


states_map <- map_data("state") #%>%
  #rename("state" = region) 
```


# Plotting map
```{r}
# colors taken from DuBois plot
# red: #DF2645
# yellow: #FAB500
# green: #4C6C5A

state_data  %>%
  ggplot(aes(fill = terciles)) +
  geom_map(aes(map_id = state), map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  #scale_fill_gradient(low = "#DF2645", high = "#FAB500", na.value = "light grey") + ## heatmap option
  scale_fill_manual(values = c("#DF2645","#FAB500","#4C6C5A"),
                    labels = c("Negative","Neutral","Positive", "Web profiles NA")) +
  labs(
    title = "Sentiment analysis of Tuskegee Airmen web profiles",
    subtitle = "Map of average sentiment scores by state",
    caption = "Figure by: Sean Bock\nData: Commemorative Air Force\n Git: https://github.com/Sdbock",
    fill = "Sentiment score:",
    x = NULL,
    y = NULL
  ) +
  theme_dubois() + ## adding dubois theme
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.border = element_blank(),
    plot.title = element_text(size = 25),
    plot.subtitle = element_text(size = 20, vjust = 10),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 15),
    plot.caption = element_text(size = 15)
  ) 


```

# Saving plot

```{r}
# saving plot in a "figures" folder
ggsave(
  filename = "figures/tidytuesday_2022_02_08.png",
  device = "png",
  width = 11,
  height = 8.5)

```
